---
title: "R Notebook"
output: html_notebook
---
# Goals 1:
Visualize volume of natural gas delivered to consumers (residential, commercial, industrial, Vehicle Fuel, Electric Power) in Massachusetts from 2015 to 2019, then find out the trend over years and which type of consumers is the biggest or smallest natural gas user. last but not least, compare with California and the US average.
# Goals 2:
Visualize Average price of natural gas delivered to consumers (residential, commercial, industrial, Electric Power) in Massachusetts from 2015 to 2019, then make a comparison with The new England region and the US average.


```{r}
library(ggplot2) 
library(plyr)
library(dplyr)
library(gridExtra)
library(scales)
library(magrittr) 
library(tidyr)

df<-read.csv('Summary Statistics for Natural Gas - Massachusetts, 2015-2019.csv', na.strings = "", check.names = FALSE) # goal 2 MA step 1
names(df)[1] <- "STATE_CODE"
names(df)[2] <- "Var"

temp_df<- filter(df, STATE_CODE =="MA")
temp_df_0<- tail(temp_df,n = 4)
temp_df_0<- select(temp_df_0, -STATE_CODE)

df_longer<-temp_df_0 %>%
  pivot_longer(-Var, 
               names_to = "years", 
               values_to = "dollars",
               values_drop_na = TRUE)
df_longer$dollars<-as.numeric(df_longer$dollars)
p1<- ggplot(data=df_longer, aes(x=years, y=dollars, fill=Var)) +
  geom_bar(stat="identity", position="dodge")+
  labs(x="",y="dollars per thousand cubic feet",title = " Average price of natural gas delivered to consumers,MA,2015‐2019" )+
  theme(plot.title = element_text(hjust=0.5,size = 15),axis.title = element_text(size = 10),legend.title=element_blank(),legend.position = "bottom",legend.text = element_text(size = 10))
p1

```

```{r}
temp_df_1 <- temp_df[33:37,,drop=F]                                             # goal 1 MA step 1
temp_df_1<- select(temp_df_1, -STATE_CODE)

df_longer_1<-temp_df_1 %>%
  pivot_longer(-Var, 
               names_to = "years", 
               values_to = "million cubic feet",
               values_drop_na = TRUE)
df_longer_1$`million cubic feet`<-as.numeric(df_longer_1$`million cubic feet`)
p2<- ggplot(data=df_longer_1, aes(x=years, y=`million cubic feet`, fill=Var)) +
  geom_bar(stat="identity", position="dodge")+
  labs(x="",y="million cubic feet",title = "Volume of Natural gas delivered to consumers,MA,2015‐2019" )+
  theme(plot.title = element_text(hjust=0.5,size = 15),axis.title = element_text(size = 10),legend.title=element_blank(),legend.position = "bottom",legend.text = element_text(size = 10))
p2+scale_y_continuous(breaks =seq(0, 150000, by = 12500))

```

```{r}
df<-read.csv('Summary Statistics for Natural Gas - CA, 2015-2019.csv', na.strings = "", check.names = FALSE) # goal 1 CA step 2 
names(df)[1] <- "STATE_CODE"
names(df)[2] <- "Var"

temp_df<- filter(df, STATE_CODE =="CA")
temp_df_1 <- temp_df[33:37,,drop=F]
temp_df_1<- select(temp_df_1, -STATE_CODE)
temp_df_1$`2015`<- as.numeric(temp_df_1$`2015`)
temp_df_1$`2016`<- as.numeric(temp_df_1$`2016`)
df_longer_1<-temp_df_1 %>%
  pivot_longer(-Var, 
               names_to = "years", 
               values_to = "million cubic feet",
               values_drop_na = TRUE)
df_longer_1$`million cubic feet`<-as.numeric(df_longer_1$`million cubic feet`)
p2<- ggplot(data=df_longer_1, aes(x=years, y=`million cubic feet`, fill=Var)) +
  geom_bar(stat="identity", position="dodge")+
  labs(x="",y="million cubic feet",title = "Volume of Natural gas delivered to consumers, CA,2015‐2019" )+
  theme(plot.title = element_text(hjust=0.5,size = 15),axis.title = element_text(size = 10),legend.title=element_blank(),legend.position = "bottom",legend.text = element_text(size = 10))
p2+scale_y_continuous(breaks =seq(0, 800000, by = 200000))

```
```{r}
df<-read.csv('Summary Statistics for Natural Gas - CA, 2015-2019.csv', na.strings = "", check.names = FALSE)# goal 1 CA MA step 3->reason: quantity, major 
names(df)[1] <- "STATE_CODE"
names(df)[2] <- "Var"

temp_df<- filter(df, STATE_CODE =="CA")
temp_df_2<-temp_df[35,,drop=F]
temp_df_2<- select(temp_df_2, -STATE_CODE)
temp_df_2$`2015`<- as.numeric(temp_df_2$`2015`)
temp_df_2$`2016`<- as.numeric(temp_df_2$`2016`)
temp_df_2$`2017`<- as.numeric(temp_df_2$`2017`)
temp_df_2$`2018`<- as.numeric(temp_df_2$`2018`)
temp_df_2$`2019`<- as.numeric(temp_df_2$`2019`)
df_longer_2<-temp_df_2 %>%
  pivot_longer(-Var, 
               names_to = "years", 
               values_to = "CA",
               values_drop_na = TRUE)

df<-read.csv('Summary Statistics for Natural Gas - Massachusetts, 2015-2019.csv', na.strings = "", check.names = FALSE)
names(df)[1] <- "STATE_CODE"
names(df)[2] <- "Var"

temp_df<- filter(df, STATE_CODE =="MA")
temp_df_3<-temp_df[35,,drop=F]
temp_df_3<- select(temp_df_3, -STATE_CODE)
temp_df_3$`2015`<- as.numeric(temp_df_3$`2015`)
temp_df_3$`2016`<- as.numeric(temp_df_3$`2016`)
temp_df_3$`2017`<- as.numeric(temp_df_3$`2017`)
temp_df_3$`2018`<- as.numeric(temp_df_3$`2018`)
temp_df_3$`2019`<- as.numeric(temp_df_3$`2019`)
df_longer_3<-temp_df_3 %>%
  pivot_longer(-Var, 
               names_to = "years", 
               values_to = "MA",
               values_drop_na = TRUE)

df_longer_2<-left_join(df_longer_2, df_longer_3, by=c("years","Var"))
df_longer_2 %>% 
   select(-Var) %>% 
   gather(State, count,CA:MA) %>% 
   ggplot(., aes(x=years, y=count, fill=State)) +
   geom_bar(stat="identity", position = "dodge")+
  labs(x="",y="count",title = "Volume of Natural gas delivered to industry, 2015‐2019" )+
  theme(plot.title = element_text(hjust=0.5,size = 15),axis.title = element_text(size = 10),legend.title=element_blank(),legend.position = "bottom",legend.text = element_text(size = 10))
```

```{r}
df<-read.csv('Summary Statistics for Natural Gas - CA, 2015-2019.csv', na.strings = "", check.names = FALSE) # goal 1 step 4 quantity
names(df)[1] <- "STATE_CODE"
names(df)[2] <- "Var"

temp_df<- filter(df, STATE_CODE =="CA")
temp_df_2<-temp_df[45,,drop=F]
temp_df_2<- select(temp_df_2, -STATE_CODE)
temp_df_2$`2015`<- as.numeric(temp_df_2$`2015`)
temp_df_2$`2016`<- as.numeric(temp_df_2$`2016`)
temp_df_2$`2017`<- as.numeric(temp_df_2$`2017`)
temp_df_2$`2018`<- as.numeric(temp_df_2$`2018`)
temp_df_2$`2019`<- as.numeric(temp_df_2$`2019`)
df_longer_2<-temp_df_2 %>%
  pivot_longer(-Var, 
               names_to = "years", 
               values_to = "CA",
               values_drop_na = TRUE)

df<-read.csv('Summary Statistics for Natural Gas - Massachusetts, 2015-2019.csv', na.strings = "", check.names = FALSE)
names(df)[1] <- "STATE_CODE"
names(df)[2] <- "Var"

temp_df<- filter(df, STATE_CODE =="MA")
temp_df_3<-temp_df[45,,drop=F]
temp_df_3<- select(temp_df_3, -STATE_CODE)
temp_df_3$`2015`<- as.numeric(temp_df_3$`2015`)
temp_df_3$`2016`<- as.numeric(temp_df_3$`2016`)
temp_df_3$`2017`<- as.numeric(temp_df_3$`2017`)
temp_df_3$`2018`<- as.numeric(temp_df_3$`2018`)
temp_df_3$`2019`<- as.numeric(temp_df_3$`2019`)
df_longer_3<-temp_df_3 %>%
  pivot_longer(-Var, 
               names_to = "years", 
               values_to = "MA",
               values_drop_na = TRUE)

df_longer_2<-left_join(df_longer_2, df_longer_3, by=c("years","Var"))
df_longer_2 %>% 
   select(-Var) %>% 
   gather(State, count,CA:MA) %>% 
   ggplot(., aes(x=years, y=count, fill=State)) +
   geom_bar(stat="identity", position = "dodge")+
  labs(x="",y="count",title = "Number of industry, 2015‐2019" )+
  theme(plot.title = element_text(hjust=0.5,size = 15),axis.title = element_text(size = 10),legend.title=element_blank(),legend.position = "bottom",legend.text = element_text(size = 10))
```

```{r}
df<-read.csv('Summary Statistics for Natural Gas - CA, 2015-2019.csv', na.strings = "", check.names = FALSE) # goal 1 step 4 major
names(df)[1] <- "STATE_CODE"
names(df)[2] <- "Var"

temp_df<- filter(df, STATE_CODE =="CA")
temp_df_2<-temp_df[47,,drop=F]
temp_df_2<- select(temp_df_2, -STATE_CODE)
temp_df_2$`2015`<- as.numeric(temp_df_2$`2015`)
temp_df_2$`2016`<- as.numeric(temp_df_2$`2016`)
temp_df_2$`2017`<- as.numeric(temp_df_2$`2017`)
temp_df_2$`2018`<- as.numeric(temp_df_2$`2018`)
temp_df_2$`2019`<- as.numeric(temp_df_2$`2019`)
df_longer_2<-temp_df_2 %>%
  pivot_longer(-Var, 
               names_to = "years", 
               values_to = "CA",
               values_drop_na = TRUE)

df<-read.csv('Summary Statistics for Natural Gas - Massachusetts, 2015-2019.csv', na.strings = "", check.names = FALSE)
names(df)[1] <- "STATE_CODE"
names(df)[2] <- "Var"

temp_df<- filter(df, STATE_CODE =="MA")
temp_df_3<-temp_df[47,,drop=F]
temp_df_3<- select(temp_df_3, -STATE_CODE)
temp_df_3$`2015`<- as.numeric(temp_df_3$`2015`)
temp_df_3$`2016`<- as.numeric(temp_df_3$`2016`)
temp_df_3$`2017`<- as.numeric(temp_df_3$`2017`)
temp_df_3$`2018`<- as.numeric(temp_df_3$`2018`)
temp_df_3$`2019`<- as.numeric(temp_df_3$`2019`)
df_longer_3<-temp_df_3 %>%
  pivot_longer(-Var, 
               names_to = "years", 
               values_to = "MA",
               values_drop_na = TRUE)

df_longer_2<-left_join(df_longer_2, df_longer_3, by=c("years","Var"))
df_longer_2 %>% 
   select(-Var) %>% 
   gather(State, count,CA:MA) %>% 
   ggplot(., aes(x=years, y=count, fill=State)) +
   geom_bar(stat="identity", position = "dodge")+
  labs(x="",y="thousand cubic feet",title = "Average Annual Consumption per industry, 2015‐2019" )+
  theme(plot.title = element_text(hjust=0.5,size = 15),axis.title = element_text(size = 10),legend.title=element_blank(),legend.position = "bottom",legend.text = element_text(size = 10))
```

```{r}
df<-read.csv('Summary Statistics for Natural Gas - Massachusetts, 2015-2019.csv', na.strings = "", check.names = FALSE) # goal 2 New England region step 2
names(df)[1] <- "STATE_CODE"
names(df)[2] <- "Var"

temp_df<- filter(df, STATE_CODE =="MA")
temp_df_1<-temp_df[51,,drop=F]
temp_df_1<- select(temp_df_1, -STATE_CODE)
df_longer_1<-temp_df_1 %>%
  pivot_longer(-Var, 
               names_to = "years", 
               values_to = "MA",
               values_drop_na = TRUE)
df<-read.csv('Summary Statistics for Natural Gas - ME, 2015-2019.csv', na.strings = "", check.names = FALSE)
names(df)[1] <- "STATE_CODE"
names(df)[2] <- "Var"

temp_df<- filter(df, STATE_CODE =="ME")
temp_df_2<-temp_df[51,,drop=F]
temp_df_2<- select(temp_df_2, -STATE_CODE)
df_longer_2<-temp_df_2 %>%
  pivot_longer(-Var, 
               names_to = "years", 
               values_to = "ME",
               values_drop_na = TRUE)
df<-read.csv('Summary Statistics for Natural Gas - CT, 2015-2019.csv', na.strings = "", check.names = FALSE)
names(df)[1] <- "STATE_CODE"
names(df)[2] <- "Var"

temp_df<- filter(df, STATE_CODE =="CT")
temp_df_3<-temp_df[51,,drop=F]
temp_df_3<- select(temp_df_3, -STATE_CODE)
df_longer_3<-temp_df_3 %>%
  pivot_longer(-Var, 
               names_to = "years", 
               values_to = "CT",
               values_drop_na = TRUE)
df<-read.csv('Summary Statistics for Natural Gas - NH, 2015-2019.csv', na.strings = "", check.names = FALSE)
names(df)[1] <- "STATE_CODE"
names(df)[2] <- "Var"

temp_df<- filter(df, STATE_CODE =="NH")
temp_df_4<-temp_df[51,,drop=F]
temp_df_4<- select(temp_df_4, -STATE_CODE)
df_longer_4<-temp_df_4 %>%
  pivot_longer(-Var, 
               names_to = "years", 
               values_to = "NH",
               values_drop_na = TRUE)
df<-read.csv('Summary Statistics for Natural Gas - VT, 2015-2019.csv', na.strings = "", check.names = FALSE)
names(df)[1] <- "STATE_CODE"
names(df)[2] <- "Var"

temp_df<- filter(df, STATE_CODE =="VT")
temp_df_5<-temp_df[51,,drop=F]
temp_df_5<- select(temp_df_5, -STATE_CODE)
df_longer_5<-temp_df_5 %>%
  pivot_longer(-Var, 
               names_to = "years", 
               values_to = "VT",
               values_drop_na = TRUE)
df<-read.csv('Summary Statistics for Natural Gas - RI, 2015-2019.csv', na.strings = "", check.names = FALSE)
names(df)[1] <- "STATE_CODE"
names(df)[2] <- "Var"

temp_df<- filter(df, STATE_CODE =="RI")
temp_df_6<-temp_df[51,,drop=F]
temp_df_6<- select(temp_df_6, -STATE_CODE)
df_longer_6<-temp_df_6 %>%
  pivot_longer(-Var, 
               names_to = "years", 
               values_to = "RI",
               values_drop_na = TRUE)
df_longer_1<-left_join(df_longer_1, df_longer_2, by=c("years","Var"))
df_longer_1<-left_join(df_longer_1, df_longer_3, by=c("years","Var"))
df_longer_1<-left_join(df_longer_1, df_longer_4, by=c("years","Var"))
df_longer_1<-left_join(df_longer_1, df_longer_5, by=c("years","Var"))
df_longer_1<-left_join(df_longer_1, df_longer_6, by=c("years","Var"))
df_longer_1 %>% 
   select(-Var) %>% 
   gather(State, count,MA:RI) %>% 
   ggplot(., aes(x=years, y=count, group = State, color =State)) +
  geom_line(size=1)+ 
  geom_point(size=2)+
  labs(x="",y="dollars per thousand cubic feet",title = "Average price of residential natural gas in New England region,2015-2019" )+
  theme(plot.title = element_text(hjust=0.5,size = 15),axis.title = element_text(size = 10),legend.title=element_blank(),legend.position = "bottom",legend.text = element_text(size = 10))

```

```{r}
df<-read.csv('Summary Statistics for Natural Gas - Massachusetts, 2015-2019.csv', na.strings = "", check.names = FALSE) # goal 2 US step 3
names(df)[1] <- "STATE_CODE"
names(df)[2] <- "Var"

temp_df<- filter(df, STATE_CODE =="MA")
temp_df_1<-temp_df[51,,drop=F]
temp_df_1<- select(temp_df_1, -STATE_CODE)
df_longer_1<-temp_df_1 %>%
  pivot_longer(-Var, 
               names_to = "years", 
               values_to = "MA",
               values_drop_na = TRUE)

df_longer_1 <- select(df_longer_1, -Var)
U_S_Price_of_Natural_Gas_Delivered_to_Residential$`Back to Contents` <- as.character(substring(U_S_Price_of_Natural_Gas_Delivered_to_Residential$`Back to Contents`,1,4)) #import file 'U_S_Price_of_Natural_Gas_Delivered_to_Residential'
df_longer_1 <- left_join(df_longer_1,U_S_Price_of_Natural_Gas_Delivered_to_Residential,by=c("years"="Back to Contents"))
colnames(df_longer_1)[3] = "US"
df_longer_1 %>% 
   gather(State, count,MA:US) %>% 
   ggplot(., aes(x=years, y=count, group = State, color =State)) +
  geom_line(size=1)+ 
  geom_point(size=2)+
  labs(x="",y="dollars per thousand cubic feet",title = "Average price of residential natural gas comparation between MA and US,2015-2019" )+
  theme(plot.title = element_text(hjust=0.5,size = 15),axis.title = element_text(size = 10),legend.title=element_blank(),legend.position = "bottom",legend.text = element_text(size = 10))
```

```{r}
df<-read.csv('Summary Statistics for Natural Gas - Massachusetts, 2015-2019.csv', na.strings = "", check.names = FALSE) # goal 1 US CA step 5
names(df)[1] <- "STATE_CODE"
names(df)[2] <- "Var"

temp_df<- filter(df, STATE_CODE =="MA")
temp_df_1<-temp_df[35,,drop=F]
temp_df_1<- select(temp_df_1, -STATE_CODE)
df_longer_1<-temp_df_1 %>%
  pivot_longer(-Var, 
               names_to = "years", 
               values_to = "MA",
               values_drop_na = TRUE)

df_longer_1 <- select(df_longer_1, -Var)
df<-read.csv('Summary Statistics for Natural Gas - CA, 2015-2019.csv', na.strings = "", check.names = FALSE) 
names(df)[1] <- "STATE_CODE"
names(df)[2] <- "Var"

temp_df<- filter(df, STATE_CODE =="CA")
temp_df_2<-temp_df[35,,drop=F]
temp_df_2<- select(temp_df_2, -STATE_CODE)
temp_df_2$`2015`<- as.numeric(temp_df_2$`2015`)
temp_df_2$`2016`<- as.numeric(temp_df_2$`2016`)
temp_df_2$`2017`<- as.numeric(temp_df_2$`2017`)
temp_df_2$`2018`<- as.numeric(temp_df_2$`2018`)
temp_df_2$`2019`<- as.numeric(temp_df_2$`2019`)
df_longer_2<-temp_df_2 %>%
  pivot_longer(-Var, 
               names_to = "years", 
               values_to = "CA",
               values_drop_na = TRUE)

df_longer_2 <- select(df_longer_2, -Var)
US_total_industry$`Back to Contents` <- as.character(substring(US_total_industry$`Back to Contents`,1,4)) #import file 'US_total_industry'
df_longer_1 <- left_join(df_longer_1,US_total_industry,by=c("years"="Back to Contents"))
df_longer_1 <- left_join(df_longer_1,df_longer_2, by="years")
colnames(df_longer_1)[3] = "US"
df_longer_1$MA<- as.numeric(df_longer_1$MA)
df_longer_1$US<-round((df_longer_1$US)/52)
df_longer_1 %>% 
   gather(State, count,MA:CA) %>% 
   ggplot(., aes(x=years, y=count, group = State, color =State)) +
  geom_line(size=1)+ 
  geom_point(size=2)+
  labs(x="",y="million cubic feet",title = "Volume of Natural gas delivered to industry, 2015‐2019" )+
  theme(plot.title = element_text(hjust=0.5,size = 15),axis.title = element_text(size = 10),legend.title=element_blank(),legend.position = "bottom",legend.text = element_text(size = 10))+scale_y_continuous(breaks =seq(44000,800000,by=100000))

```
```{r}
library("readxl")
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
gas_price <- read_excel("gasdata.xls",sheet = 'Data 1')

imports_exports <- read_excel("gasdata.xls",sheet = 'Data 2')

underground_storage <- read_excel("gasdata.xls",sheet = 'Data 3') #skipped due to not enough data

liquefied_storage <- read_excel("gasdata.xls",sheet = 'Data 4')

consumption <- read_excel("gasdata.xls",sheet = 'Data 5')

prices <- gas_price[2]
prices <- prices %>%
  drop_na()

imports <- imports_exports[2]

net_withdrawals <- liquefied_storage[4]

residential_consumption <- consumption[5]


p <- prices %>% slice(3:33)  # price 1989 - 2019

i <- imports %>% slice(9:39) # import 1989 - 2019

w <- net_withdrawals %>% slice(12:42) # net withdrawal 1989 - 2019

c <- residential_consumption %>% slice(25:55) # residential_consumption 1989 - 2019


names(p)[1] <- "price"
names(i)[1] <- "import"
names(w)[1] <- "net withdrawal"
names(c)[1] <- "resitent_consumption"

# Organized dataframe with all information needed during year 1989 - 2019

df <- data.frame(year = c(1989:2019),
                price=as.numeric(as.character(p$price)),
                import=as.numeric(as.character(i$import)),
                withdrawal=as.numeric(as.character(w$`net withdrawal`)),
                consumption=as.numeric(as.character(c$`resitent_consumption`)))

# Plot 1 gas price

plot1 <- ggplot(data = df, aes(x = year, y = price)) + 
  geom_bar(stat='identity', fill = 'light blue') +
  labs(title = "Gas price over year 1989-2019(MA)") +
  theme_bw()

# Plot 2 gas import

plot2 <- ggplot(data = df, aes(x = year, y = import)) + 
  geom_line(color = 'red',size = 1.5) + 
  labs(title = "Import level over year 1989-2019(MA)") +
  theme_bw()

# Plot 3 gas net withdrawal

plot3 <- ggplot(data = df, aes(x = year, y = withdrawal)) + 
  geom_bar(stat='identity', fill = 'red') +
  labs(title = "Net withdrawal level over year 1989-2019(MA)") +
  theme_bw()

# Plot 4 gas consumption

plot4 <- ggplot(data = df, aes(x = year, y = consumption)) + 
  geom_line(color = 'dark blue', size = 1.5) +
  labs(title = "Consumption over year 1989-2019(MA)") + 
  theme_bw()


grid.arrange(plot1)
grid.arrange(plot2)
grid.arrange(plot3)
grid.arrange(plot4)


grid.arrange(plot1, plot2, plot3, plot4, ncol = 2, heights=c(1,1))


```
Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
